Supplementary material of the article "Are historical stage records useful to decrease the uncertainty of flood frequency analysis ? A 200-year long case study" from Journal of Hydrology

### 2023-06-14 ArticleChProp/SupplementaryMaterial.tex
second submission

%bib options
\usepackage[backend=biber,style=authoryear,bibstyle=authoryear,natbib=true,
giveninits=true,uniquename=false,uniquelist=false,maxcitenames=2,date=year,
maxbibnames=99,url=false]{biblatex}
\addbibresource{Article.bib}
\usepackage{dsfont}
\usepackage{multirow}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{subcaption}
\usepackage{authblk}
%hyperlinks options
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,filecolor=magenta, urlcolor=cyan,citecolor=cyan}
\geometry{a4paper, total={180mm,257mm}, left=20mm, top=20mm}
%double spacing and line numbers for submission
\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\doublespacing

\title{Are historical stage records useful to decrease the uncertainty of flood frequency analysis ? A 200-year long case study}

@@

\begin{abstract}

    Flood frequency analysis (FFA), a widely used method to estimate flood hazard, is affected by several sources of uncertainty. Extending flood samples by reanalyzing historical continuous stage records has the potential to reduce sampling uncertainty, but the historical flood discharges derived from this reanalysis are generally affected by large uncertainties. This paper explores whether historical stage records improve design flood estimates through a chain of uncertainty estimation methods for FFA. Uncertainties are estimated and propagated from stage and rating curves to design flood estimates using Monte Carlo procedures. The role of both streamflow and sampling uncertainties in design flood estimation is examined. This procedure  is applied to the 205-year long continuous stage series of the Rhône River at Beaucaire, France (95 590 km\textsuperscript{2}). The estimated streamflow 95\% uncertainty varies from 30\% (XIX\textsuperscript{th} Century) to 5\% (1967-2020). The total uncertainty of design flood is significantly reduced when the length of the series increases from 20 to 100 years due to sampling uncertainty reduction. However, the total uncertainty remains stable beyond this sample size: this is because large uncertainties affecting the XIX\textsuperscript{th} Century flood discharges compensate for the reduction in sampling uncertainty. Enlarging the sample size to two centuries leads to including the two largest known floods in 1840 and 1856. In turn, this induces a 15\% increase of the 1000-year flood estimates.
    \newline
    \textbf{Keywords}: Flood frequency analysis, Historical stage records, Uncertainty propagation, Streamflow uncertainty, Sampling uncertainty

\end{abstract}

\section{Introduction}
    \paragraph{}
    Flood frequency analysis (FFA) is a widely used method to estimate flood hazard. It allows linking the magnitude of a flood to its probability of occurrence (\citet{hamed_flood_2019}; \citet{jain_design_2019}). Flood estimates for various exceedance probabilities or, equivalently, return periods, are commonly used for population safety policies, land use planning, as well as industrial safety. The standard FFA approach is to estimate a distribution using a sample of flood peaks, typically defined as annual maximum discharges or discharges over a given threshold. This distribution may be extrapolated to reach the desired flood quantile that typically corresponds to a 100 or 1000-year return period (see \cite{le_delliou_recommandations_2014} for dam safety regulations in France).

    \paragraph{}
    This FFA approach is affected by various sources of uncertainty. First, the hydrological data used to estimate the FFA distribution is uncertain. Indeed, streamflow time series are generally derived from stage time series through rating curve models \citep{rantz_measurement_1982}. This procedure includes three types of errors on hydrological data: stage measurement; discharge measurement (gaugings); stage-discharge model (rating curve). Moreover, the estimated FFA distribution is also affected by sampling uncertainty resulting from the limited size of available streamflow data \citep{kjeldsen_uncertainty_2011}. Considering the importance of decisions relying on FFA results, a consistent treatment of uncertainty all over the data processing chain (including both streamflow and sampling uncertainties) is essential but is usually not performed.

    \paragraph{}
    Streamflow series are affected by several sources of uncertainty as described by \citet{mcmillan_benchmarking_2012}. First, a large number of stage error sources are identified in the literature (\citet{van_der_made_determination_1982}; \citet{petersen-overleir_uncertainty_2005}; \citet{mcmillan_benchmarking_2012}; \citet{horner_impact_2018}), such as staff gauge reading, levelling of the staff gauge, or stage sensor calibration. The frequency of measurement may also induce time interpolation errors. With modern automatic gauges, stage is generally measured with a time step small enough (e.g. between 15 min and 1 hour) to get negligible time interpolation errors. However, before the rise of automatic gauges, measurements were made by operators who read the staff gauge less frequently (e.g. once or a few times per day), thus possibly missing the flood peak. This issue is particularly critical when old stage series are used. \citet{hamilton_quantifying_2012} and \citet{kuentz_hydrometrie_2014} estimated the measurement frequency error by sub-sampling recent, sub-hourly measurements. They calculated the difference between the variable of interest (such as the daily maximum stage) derived from scarce data, and the same variable derived from high-frequency measurements. \citet{kuentz_hydrometrie_2014} applied the (monthly-averaged) calculated bias to correct old stage series. This correction aimed at taking into account the error due to the daily variability caused by snow melt. However, this type of correction has never been applied to peak stage correction during floods, especially in the case of long stage series.

    \paragraph{}
    Rating curve uncertainty is also a major issue when dealing with streamflow series. Transforming stage into discharge requires calibration data (gaugings) to establish the stage-discharge relationship. Note that the term "gauging" corresponds in this paper (and in the French practice) to the sporadic measurement of the stage-discharge couple, which can correspond to the term "discharge measurement" in the British or American practice. Gaugings uncertainty depends on the measurement method (\citet{lecoz_quantification_2014}; \citet{puechberty_charte_2017}). Moreover, the rating curve is also affected by uncertainties coming from the imperfection of the chosen model to represent the actual hydraulic configuration, and from parameter estimation. Many methods have been proposed to quantify these uncertainties (\citet{petersen-overleir_bayesian_2009}; \citet{juston_rating_2014}; \citet{le_coz_combining_2014}; \citet{morlot_dynamic_2014}; \citet{coxon_novel_2015}; \citet{mcmillan_rating_2015}; \citet{mansanarez_rapid_2019}). A comparison of several of these methods has been recently proposed by \citet{kiang_comparison_2018}. Another important issue affecting streamflow data accuracy is rating changes. The stage-discharge relationship is frequently affected by changes caused by various factors, either natural or anthropic, for instance: bed geometry evolution during floods or river works, aquatic vegetation growth and decay, ice cover... A regular monitoring through gaugings is essential to detect those changes \citep{ibbitt_gauging_1987} that can be transient or sudden. Several methods have been proposed to deal with rating changes: estimating rating curves on moving temporal windows (\citet{westerberg_stage-discharge_2011}; \citet{guerrero_temporal_2012}), computing as many rating curves as there are gaugings \citep{morlot_dynamic_2014}, exploring changes in the annual minimum stages \citep{lapuszek_methods_2015}, selecting the 0.5-year return period discharge as a threshold for rating changes \citep{mcmillan_impacts_2010}. More recently, \citet{darienzo_detection_2021} proposed a method based on a recursive segmentation procedure, accounting for both gaugings and rating curve uncertainties. This method has a particular interest when dealing with old and uncertain gaugings. Following the detection of rating shifts, rating curves should be estimated for each stable period. This task may not be straightforward, as the number of gaugings available within a stable period is not always sufficient to properly estimate the stage-discharge relationship for the whole discharge range. A common way to address the lack of gaugings (in particular flood gaugings) within some of the stable periods is to use gaugings from the other stable periods (\cite{mcmillan_benchmarking_2012}; \cite{puechberty_charte_2017}). \citet{mansanarez_shift_2019} proposed an alternative approach to deal with this issue. They developed a stage-period-discharge (SPD) model where rating curve parameters may vary across periods, while others are supposed constant. This method has the advantage of transferring information between periods to improve the rating curve estimation, even when few gaugings are available.

    \paragraph{}
    Estimating sampling uncertainty in FFA is a well-established approach. Whatever the chosen distribution and estimation method, standard statistical procedures are available \citep{coles_classical_2001}. However, these standard procedures only quantify sampling uncertainty, they do not consider data uncertainty. The literature review proposed in the previous paragraphs shows that methods for quantifying individual sources of uncertainty (stage, rating curve and FFA distribution estimation) are available. However, the way these multiple uncertainties propagate through the FFA analysis chain has been less thoroughly studied. A few solutions have emerged to propagate uncertainties in stage time series through uncertain rating curves (\citet{dymond_accuracy_1982}; \citet{herschy_hydrometry_1998}; \citet{petersen-overleir_uncertainty_2005}), but they assume independent stage errors and therefore neglect systematic errors. \citet{horner_impact_2018} proposed a method for the propagation of both sources of stage uncertainty through uncertain rating curves. Therefore, it is possible to distinguish the effects of independent and systematic stage errors on streamflow uncertainty. \citet{petersen-overleir_accounting_2009}, \citet{steinbakk_propagation_2016}, and \citet{vieira_assessing_2022} performed an integrated analysis in which both rating curve parameters and flood frequency distribution are estimated. These studies highlighted the importance of considering rating curve uncertainty for design flood estimations and concluded that, under some conditions, accounting for rating curve uncertainty may notably widen the uncertainty intervals around flood quantiles. However, these studies did not consider stage measurement and time interpolation errors nor rating changes, which may constitute a major source of uncertainty for streamflow data. This is particularly the case when dealing with long streamflow series for which stage uncertainty is large and variable through time, and rating changes may have been missed. Their consideration in a flood frequency framework is therefore not avoidable.

    \paragraph{}
    The following questions will be considered in this paper: 
    \begin{itemize}
        \item[1.] How to make the most of historical hydrometric (based on stage records) data in flood frequency analysis while accounting for multiple and variable uncertainties at each step of the procedure ?

        \item[2.] What is the contribution of each source of uncertainty to the flood quantile uncertainty when historical data are taken into account ?

        \item[3.] To what extent does enlarging streamflow samples by adding increasingly uncertain historical data improve flood quantiles estimation ? How are the relative contributions of sampling and streamflow uncertainties evolving with sample size ?
    \end{itemize}

    \paragraph{}
    Note that the "historical data" term used in this paper refers to the use of old but continuously measured stage series, as opposed to sporadic flood marks, prior to regular stage measurements.

    \paragraph{}
    This paper illustrates the chained application of methods to quantify and propagate uncertainty from stage records (and their limited time resolution) and stage-discharge rating curves to the estimation of flood distribution with uncertainties. While most of these methods already exist, a key novelty of this work is their combination in a consistent framework (Figure \ref{fig:ChProp}) to provide an end-to-end evaluation of the  uncertainty affecting FFA estimates. An original method to quantify the stage uncertainty stemming from infrequent readings is also proposed.

    The paper is organized as follows. First, the methodology for establishing uncertain streamflow series in a century-long context is presented. It goes through the detection of rating shifts (section \ref{subsec:RatingShifts}), the estimation of rating curves (section \ref{subsec:RC_SPD}), and the estimation (section \ref{subsec:StageErr}) and propagation (section \ref{subsec:PropagStage}) of stage errors. Then, an approach to propagate streamflow uncertainty through the estimation of extreme flood quantiles is proposed (section \ref{subsec:FFA}). This procedure is applied to the Beaucaire gauge on the Rhône River (section \ref{sec:Bcr}), which previous official FFA only used a 80-year long discharge series \citep{rigaudiere_etude_2000}. In France, the official design flood for flood risk mapping is based on the “largest known flood” or the Q100 flood (AEP=0.01) if the latter is greater. Uncertainty is not taken into account in the official rules. The recent works of \citet{pichard_hydro-climatology_2017} and \citet{bard_actualisation_2018} provided a continuous stage series from 1816 to the present time, which makes it the ideal case study for demonstrating this procedure. The results of this application are presented in section \ref{sec:Results}, and they are discussed in section \ref{sec:Discussion}, where avenues for improvements are proposed.

    \begin{figure}[h!]
    \centering
        \includegraphics[width = 8cm]{Figs/1-uTotSchema.png}
        \caption{Block diagram of the uncertainty propagation procedure. Grey blocks represent data, blue blocks stand for analysis methods/models that correspond to the sub-sections of this article. $h$ is the water stage, $uh$ the stage uncertainty, $Q$ the discharge, $uQ$ the discharge uncertainty, $t$ is the time and $RP$ the return period of flood quantiles.}
        \label{fig:ChProp}
    \end{figure}
    \FloatBarrier

\section{Uncertainty propagation chain for flood frequency analysis}
     \subsection{Rating shifts detection}
     \label{subsec:RatingShifts}

    The stage-discharge relationship is sensitive to sudden changes caused by morphogenic floods or other causes affecting the flow characteristics. Relying on residuals between the gaugings and the rating curve is the most common approach to monitor the stability of this relationship over time. The method proposed by \citet{darienzo_detection_2021} is used in this work and can be summarized as follows. First, a baseline rating curve is estimated from the whole gaugings dataset. The residuals between gaugings and the rating curve are determined, and a statistical segmentation procedure is applied to them. This procedure accounts for the residuals uncertainty, coming from both the gaugings uncertainty and the rating curve uncertainty. The optimal number of stable sub-periods is determined based on the Bayesian Information criterion (BIC). Then, the same steps are applied recursively to each sub-period. The recursive procedure is stopped when the BIC indicates that a single period is optimal for all sub-periods. The results are not only the dates of the rating shifts but the posterior probability density functions (pdf) of change point times. This allows affecting the shift time to the time of the maximum stage included in the posterior 95\% credibility interval. Prior knowledge is provided on the mean of the residuals in each sub-period. The maximum number of segments at each iteration also needs to be specified. All technical details can be found in \citet{darienzo_detection_2021}.

    \subsection{Multi-period rating curves estimation: stage-period-discharge model}
    \label{subsec:RC_SPD}
    \paragraph{}
    Once the stable periods have been identified, the next step is to estimate the rating curves. \citet{mansanarez_shift_2019} developed a stage-period-discharge (SPD) model "based on the physical interpretation of changes in the stage-discharge relation across a series of stability periods". The SPD model is based on the BaRatin model \citep{le_coz_combining_2014}.
    \paragraph{} BaRatin uses the Bayesian paradigm to estimate the parameters of the rating curve equation, a combination of power equations: $Q = a(h-b)^c$, where $Q$ is the discharge, $h$ is the stage, $b$ is an offset (corresponding to the cease-to-flow stage), and $a$ and $c$ are the coefficient and the exponent of the power function. The rating curve equation is deduced from a hydraulic analysis of the gauging station, aimed at identifying the main hydraulic controls governing the stage-discharge relation. The multiple controls can be activated successively or simultaneously. Bayesian inference allows deriving the posterior distribution of rating curve parameters by combining hydraulic information (priors for parameters of each hydraulic control) and information from gaugings with uncertainty (likelihood). Two sources of uncertainty are associated with the estimated rating curve. Parametric uncertainty reflects the uncertainty due to the rating curve parameters estimation because of the limited amount of gaugings and the gaugings uncertainty. Remnant uncertainty comes from the imperfection of the chosen rating curve model to represent the actual hydraulic configuration. The posterior distribution is explored using a Markov Chain Monte Carlo (MCMC) sampler, leading to $m$ realizations of the rating curve parameter vector representing parametric uncertainty. We refer the reader to \citet{le_coz_combining_2014} for a more thorough description.

    \paragraph{}
    The SPD model estimates the rating curves of each stable period based on the same principle by considering that some parameters vary in time, while others remain constant throughout the stable periods. An important step is the identification of those varying parameters based on an hydraulic analysis of the site. Generally, channel depths and/or widths are suspected to change. A distinction is made between "local changes" affecting the lowest control only (for instance the movement of the controlling riffle) and "overall changes" affecting several controls at the same time (for instance, the scouring or filling of the main channel, affecting the offsets of both the low-flow controlling riffle and the main channel itself). Prior specification for varying parameters can be based on the analysis of the yearly lowest stages, which provide information on the evolution of riverbed elevation, as described by \citet{lapuszek_methods_2015}. See \citet{mansanarez_shift_2019} for a detailed description of prior specification for time-varying rating curves. Specific investigation is needed to correctly account for past flood surveys of the cross section, as complex filling/erosion process may have been encountered during a flood with morphogenic changes.

    \subsection{Stage uncertainties}
    \label{subsec:StageErr}

    Many sources of error having distinct statistical properties can affect stage measurements, as described in \citet{horner_impact_2018}. Five different sources of error ($\delta_{1,...,5}$) affecting stage measurements are considered. Let $h(t)$ be the measured maximum stage of a day $t$. The unknown true maximum stage $\hbar(t)$ is assumed to be approximated by the following equation:

    \begin{equation}
        \hbar(t) = h(t) + \delta_1(t) + \delta_2(t) + \delta_3(t) + \delta_4(t) + \delta_5(t)
        \label{eq:StageError}
    \end{equation}

    \paragraph{}Staff gauge reading errors $\delta_1 \sim \mathcal{N}(0,\sigma_1)$ originate from operators reading the gauge, where $\sigma_1$ depends on the resolution of the graduations (usually 1 cm), and can be increased by waves, especially during floods \citep{mcmillan_benchmarking_2012}.

    \paragraph{}Nowadays, most stage measurements are done with automatic sensors of various types such as pressure sensors, floats, radars, and they require a calibration to link the water stage to the measured proxy (respectively the pressure of the water column, the height of a float, or the air draught). Two types of errors arise from this process: sensor errors $\delta_2 \sim \mathcal{N}(0,\sigma_2)$, where $\sigma_2$ is usually estimated by the sensor manufacturer, and sensor calibration errors $\delta_3 \sim \mathcal{N}(0,\sigma_3)$ that are related to the corrections made by operators when comparing the stage measured by the sensor to the actual stage at the staff gauge reference. An operator error at this step could affect the stage measurement until the next calibration. Sensor calibration error $\delta_3$ is hence assumed constant between two calibrations and can be represented by drawing a new random value at each operator intervention.

    \paragraph{}Datum errors $\delta_4  \sim \mathcal{N}(0,\sigma_4)$ are related to changes in the datum reference elevation of the staff gauge zero value and possible discontinuity between successive gauges. Similarly to $\delta_3$, this error is constant between two gauge changes or datum reference measurements.

    \paragraph{}Measurement frequency errors $\delta_5$ are related to the inadequacy of the frequency of measurement with respect to the rate of stage variations, leading to the true daily maximum occurring in between measurements. Unlike other types of stage errors, this error is hence necessarily positive, which calls for using a positive distribution such as the Exponential distribution. The parameters of this distribution can be estimated with data from the recent period, by analyzing the difference between the daily maximum stage derived from the high-frequency sensor measurement and that from an infrequent fixed-time reading. Note that the frequency errors for hourly (or more frequent) measurements are considered negligible for large rivers with slow variations, such as the Rhône River at Beaucaire.

    \paragraph{}
    To sum up, $\delta_1$, $\delta_2$ and $\delta_5$ errors are drawn at each measurement time step, while $\delta_3$ and $\delta_4$ errors are only drawn at specific calibration times. Errors $\delta_1$ to $\delta_4$ are assumed Gaussian with known standard deviations, while $\delta_5$ is assumed Exponential with parameter estimated by subsampling recent measurements. For each error type, 500 realizations are drawn from their respective distribution. Applying eq. \ref{eq:StageError}, the total stage uncertainty is therefore represented by 500 possible realizations of the stage $\hbar(t)$.

   \subsection{Propagation of stage and rating curve uncertainties to streamflow time series}
   \label{subsec:PropagStage}
   Stage realizations can be propagated through uncertain rating curves, following the approach described by \citet{horner_impact_2018}. Four cases are considered to estimate the contributions of the different sources of streamflow uncertainty:

   \begin{itemize}
       \item \textbf{Case 1: Maxpost streamflow.} Stage is taken as the median of the stage time series realizations. This unique stage time series is propagated through the maxpost (Maximum A Posteriori: obtained with parameters maximizing the posterior pdf) rating curve, resulting in a single discharge time series.

       \item \textbf{Case 2: Stage uncertainty.} The $n$=500 possible stage time series are propagated through the maxpost rating curve. Thus, $n$ discharge time series are obtained.

       \item \textbf{Case 3: Stage and parametric rating curve uncertainty.} The $n$=500 realizations of stage time series are propagated through $m$ rating curves, corresponding to the $m$ MCMC-simulated parameter vectors described in section \ref{subsec:RC_SPD}. This leads to $n \times m$ discharge time series.

       \item \textbf{Case 4: Total streamflow uncertainty.} It is obtained by adding remnant rating curve uncertainty (as defined in section \ref{subsec:RC_SPD}) to case 3. To achieve this, $n \times m$ time series of remnant errors are sampled from their estimated distribution and added to the time series created for case 3. 
   \end{itemize}

    \subsection{Estimation of probabilistic model parameters and flood frequency analysis}
    \label{subsec:FFA}

    The Generalized Extreme Value (GEV) distribution is commonly used to model annual maximum discharges (AMAX) (see \citet{hamed_flood_2019} or \citet{jain_design_2019}). The vector $\boldsymbol{\theta} = (\mu,\sigma,\xi)$ denotes the location, scale and shape parameters of the GEV distribution. The parameters can be estimated based on an independent and identically distributed ($iid$) sample of $j$ annual maximum discharges $(q_t)_{t=1,...,j}$. Bayesian-MCMC estimation is used in this work, as described in \citet{coles_classical_2001}. The posterior distribution quantifies sampling uncertainty and can be represented by $r$ MCMC-generated GEV parameter vectors $\boldsymbol{\Theta} = (\boldsymbol{\theta}_1,...,\boldsymbol{\theta}_r)$. The maxpost vector is noted $\boldsymbol{\hat{\theta}}$.

    \paragraph{}
    As described in section \ref{subsec:PropagStage}, total streamflow uncertainty is represented by $n \times m$ possible realizations of the streamflow, hence of the AMAX series $(q_t^{(i)})_{i=1,...,n \times m;\;t=1,...,j}$, that are subsampled to $s$=500 realizations to reduce computation time. The estimated flood quantiles should consider both sampling and streamflow uncertainties. Similarly to \citet{steinbakk_propagation_2016}, the aim is to estimate the contribution of each source to the total uncertainty. For this purpose, three cases can be considered:

    \begin{itemize}
        % MP: median of stage err x maxpost RC x maxpost GEV
        \item \textbf{Case 1: Maxpost quantiles.}        
        The GEV distribution is estimated using the  single AMAX series $\mathbf{\hat{q}} = (\hat{q_t})_{t=1,...,j}$ derived from the maxpost streamflow series (case 1 in section \ref{subsec:PropagStage}). Flood quantiles are then computed using the maxpost GEV parameters $\boldsymbol{\hat{\theta}} =  (\hat{\mu}, \hat{\sigma}, \hat{\xi})$. In this case, both streamflow and sampling uncertainties are ignored.

        % Hydro U: hydrometric U realizations x maxpost GEV
        \item \textbf{Case 2: Streamflow uncertainty}. The GEV distribution is estimated for each possible AMAX realization: $\mathbf{q}^{(i)} = (q_t^{(i)})_{t=1,...,j;\;i=1,...,s}$. However, only the maxpost GEV parameters vector is retained for each realization. This results in $s$ vectors of GEV parameters $(\boldsymbol{\hat{\theta}}^{(i)})_{i=1,...,s}$ that represent the effect of streamflow uncertainty of flood quantiles, ignoring sampling uncertainty.

        % Total U
        \item \textbf{Case 3: Total uncertainty.} Similarly to Case 2, the GEV distribution is estimated for each of the $s$ realizations of the AMAX series, but all the $r$ MCMC-simulated GEV parameters are used, leading to $s \times r$ vectors of GEV parameters $(\boldsymbol{\theta}^{(i)}_k)_{k=1,...,r;\;i=1,...,s}$. The result thus reflects both sampling and streamflow uncertainties.
    \end{itemize}

\section{Case study: The Rhône River at Beaucaire}
\label{sec:Bcr}
    \subsection{Site}

    The Rhône River at Beaucaire (95 590 km\textsuperscript{2}) is the lowest gauge of the Rhône River (Figure \ref{fig:locstations}). It captures all the complexity of the Rhône River hydrological regime, from the Alpine area to the oceanic and Mediterranean influences. The annual mean discharge is around 1700 m\textsuperscript{3}/s \citep{bard_actualisation_2018}, and the maximum known discharge reached 12 500 $m^3/s$ (May 1856, \citet{lang_les_2014}). The station lies in a flood sensitive area, as illustrated by the recent 2003 flood, resulting in 1.1 billion euros worth of damage \citep{lang_les_2014}. 
    The first stage measurements started in 1816, close to the bridge linking the cities of Beaucaire and Tarascon. This station is named "Pont de Beaucaire" (Kilometric point 267.6 from Lyon). It has been used until the construction of the Vallabrègues hydroelectric scheme in 1967, which led to the derivation of a part of the discharge. Consequently, a new gauging station was installed 2~km downstream from the original one, downstream from the restitution of the derivated discharges. This station, logically named "Beaucaire Restitution" (Kilometric point 269.5), has been used ever since. This resulted in a data gap during the construction process between 1967 and 1970.

    \begin{figure}[h!]
        \centering
        \includegraphics[width = 11cm]{Figs/2-BcrBv.png}
        \caption{The French Rhône River catchment and Beaucaire gauging stations (from www.geoportail.gouv.fr and www.openstreetmap.org)}
        \label{fig:locstations}
    \end{figure}
    % \FloatBarrier

    \subsection{Rating curves}
    \label{sec:prior_elicitation}

	\paragraph{} Many gauges of the Rhône River are subject to the effect of variable backwaters caused by the proximity of a dam, and therefore require the use of a stage-fall-discharge (SFD) rating curve model (for example, Valence gauge, 140 km upstream from Beaucaire described by \cite{mansanarez_non-unique_2016}). Beaucaire is located within a narrowing of the floodplain and there is no dam downstream from the gauge. However, a backwater effect from the sea has been observed at Beaucaire Restitution, but it only affects the very low flows. As this article focuses on floods, we assume that there is no reason to use a SFD model here. Consequently, like the gauge operator (CNR), a stage-discharge (SD) model is used for both Pont de Beaucaire and Beaucaire Restitution gauges.

        \subsubsection{Pont de Beaucaire}

         At Pont de Beaucaire, the stage-discharge relationship can be approximated by two additive channel controls: a main channel and a floodway. Thus, the rating curve equation can be written as follows:

        \begin{equation}
        Q(h) =
          \begin{cases}
           a_1(h-b_1)^{c_1}, & \text{if $\kappa_1 < h \leq \kappa_2$ (main channel) }\\
           a_1(h-b_1)^{c_1}+ a_2(h-b_2)^{c_2}, & \text{if $h > \kappa_2$ (main channel + floodway)}
          \end{cases}
          \label{eq:RcPt}
        \end{equation}

    Within the main channel (when water stage is below $\kappa_2 \approx$ 2~m), the flow is splitted in two sub-channels (figure \ref{subfig:matrixCh}) since time immemorial (at least before 1816) as described by \citet{armand_ii_1907}. The mobile sandbars separating the flow were progressively fixed by dikes during the XIX\textsuperscript{th} Century to ease the navigation (figure \ref{subfig:diguesarmand}). These sub-channels are connected upstream and downstream from the gauge location, thus they can be modelled as a single main channel whose average width ($\approx$ 300~m) is the sum of the two sub-channels widths (figure \ref{subfig:matrixCh}). When stage exceeds $\kappa_2$, water starts flowing on the sandbars between the two subchannels. At the gauge location, the total width is limited by unsubmersible levees, but a floodway is also activated a few hundred meters downstream from the station, impacting the stage-discharge relationship at the gauge. The width of this floodway is around 500~m.   
    \paragraph{}
    The prior distributions of the rating curve parameters are specified using historical material retrieved in regional archives, as described in table \ref{tab:PriorPt}. Physical parameters that have a direct hydraulic meaning are expressed in the first three lines: channel width ($B$), slope ($S$) and Strickler coefficient ($K$). The resulting prior distribution for the inferred parameter $a=KB\sqrt S$ is deduced by Monte Carlo propagation. Log-normal priors are used for positive quantities such as slopes, channel widths and Strickler coefficients. Informative but imprecise priors are assigned to parameters such as channel widths, slopes or offsets which can be difficult to estimate precisely. For $c$ exponents, very precise priors are used because they depend on the control type and shape (here $c=5/3$ for wide rectangular channel controls based on the simplified Manning-Strickler equation as described by \citet{le_coz_combining_2014}). Structural uncertainty parameters have uninformative priors.
    \paragraph{}
    According to historical profiles and cross-sections, we assume that changes affecting main channel and floodway controls may have occurred due to major floods (in particular 1840, 1856 and 1935 floods) and that channel widths remained constant. Those changes are called "overall changes" and are supposed to affect both main channel and floodway offsets ($b_1$ and $b_2$) at the same time.
    Meanwhile, we assume that local changes due to dike works or sediment depositions from small floods affected the offset ($b_1$) of the main channel only. As described by \citet{mansanarez_shift_2019}, local and overall changes $\Delta_l^{(k)}$ and $\Delta_g^{(k)}$ affect the offsets of two consecutive periods ($(k-1)$ and $k$) as follows:

       \begin{equation}
          \begin{cases}
           b_1^{(k)} = b_1^{(k-1)}-(\Delta_g^{(k)}+\Delta_l^{(k)}), & \text{(incremental changes in the main channel)}\\
           b_2^{(k)} = b_2^{(k-1)}-\Delta_g^{(k)}, & \text{(incremental changes in the floodway)}
          \end{cases}
          \label{eq:SPD_Pt}
        \end{equation}

       As the most recent period obtained by gaugings segmentation is assumed to be the most accurately known, it is used as the reference period ($k=1$) and periods are numbered backward in time. Prior distributions of offset changes are determined in section \ref{sec:stageevolution}.

        \begin{figure}[h!]
            \centering
            \begin{subfigure}{0.7\linewidth}
            \centering
            \includegraphics[width=1\linewidth]{Figs/3a-DiguesArmand.png}\hfill
            \caption{}
            \label{subfig:diguesarmand}
            \end{subfigure}

            \begin{subfigure}{0.6\linewidth}
            \centering
            \includegraphics[width=1\linewidth]{Figs/3b-MatrixChenal_EN.png}
            \caption{}
            \label{subfig:matrixCh}
            \end{subfigure}

            \begin{subfigure}{.9\linewidth}
            \includegraphics[width=.48\linewidth]{Figs/3c-ProfilsBardRestit.png}
            \includegraphics[width=.48\linewidth]{Figs/3c-ProfilAvalRestitEN.png}
            \caption{}
            \label{subfig:avalprofilesRestit}
            \end{subfigure}

            \caption{Historical geometry of the Rhône river near Beaucaire: (a) Map of dike evolution between 18\textsuperscript{th} and 20\textsuperscript{th} centuries, adapted from \citet{armand_ii_1907}; (b) Approximation of the two subchannels composing the main channel control, based on a 1845 cross-section survey; (c) Profiles from 1974 to 2016 (left) at Beaucaire Restitution station and 2.5 km downstream from the station (right) from CNR data, translated from \citet{bard_actualisation_2018} and \citet{medd_debit_2005}}
            \label{fig:groupPriorPt}
        \end{figure}
        % \FloatBarrier

        \subsubsection{Beaucaire Restitution}

        Beaucaire Restitution station has a quite stable profile according to 1974-2016 cross-sections (figure \ref{subfig:avalprofilesRestit} left), but the stage-discharge relationship is known to be influenced by the Mediterranean Sea level variations for very low flows (this influence does not apply to the Pont de Beaucaire gauge, located 2 kilometers upstream). This backwater effect can be represented by a channel control with a slope smaller than the slope of the uniform flow, i.e. the mean slope of the channel. The first control (representing low flows influenced by the sea) therefore has the same geometry as the second control (the main channel), but a smaller slope. The main channel control is not influenced by the sea and its slope is close to the longitudinal river slope. 
        \paragraph{}At the gauge location, the 12 meters high banks prevent overbank flows (figure \ref{subfig:avalprofilesRestit} left). However, overbank flows occur further downstream on the left bank, for stages higher than approximately 8 m (figure \ref{subfig:avalprofilesRestit} right). A floodway control (additive to the main channel) is activated above $\approx$ 8 m to model those overbank flows. Therefore, the rating curve equation can be written as follows:

        \begin{equation}
        Q(h) =
          \begin{cases}
           a_1(h-b_1)^{c_1}, & \text{if $\kappa_1 < h \leq \kappa_2 $ (main channel, sea-influenced) }\\
           a_2(h-b_2)^{c_2}, & \text{if $\kappa_2 < h \leq \kappa_3 $ (main channel, non-influenced)}\\
           a_2(h-b_2)^{c_2}+ a_3(h-b_3)^{c_3} & \text{if $h > \kappa_3$ (main channel + floodway)}
          \end{cases}
          \label{eq:RcRes}
        \end{equation}

        Prior distributions of rating curve parameters are specified using recent maps and cross-sections. Priors of the influenced and non-influenced main channels offsets $b_1$ and $b_2$ are assumed Gaussian with mean the riverbed elevation that is approximately equal to -5 m. Those offsets $b_1$ and $b_2$ are assumed changing in parallel (same local changes as the controls are in the same channel) due to a bed erosion trend described in section \ref{sec:stageevolution}, whereas floodway offset $b_3$ and channel widths are supposed constant because of fixed dikes. These "local changes" $\Delta_l^{(k)}$ are computed backwards in time as follows:

>>>>

%bib options
\usepackage[backend=biber,style=authoryear,bibstyle=authoryear,natbib=true,
giveninits=true,uniquename=false,uniquelist=false,maxcitenames=2,date=year,
maxbibnames=99,url=false]{biblatex}
\addbibresource{Article.bib}
\usepackage{dsfont}
\usepackage{multirow}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{subcaption}
\usepackage{authblk}
%hyperlinks options
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,filecolor=magenta, urlcolor=cyan,citecolor=cyan}
\geometry{a4paper, total={180mm,257mm}, left=20mm, top=20mm}

\title{Supplementary material of the article "Are historical stage records useful to decrease the uncertainty of flood frequency analysis ? A 200-year long case study" from Journal of Hydrology}

@@

	\begin{figure}[h]
		\centering
		\includegraphics[width=\linewidth]{Supplementary/FloodStages.pdf}
	    \caption{Stage measurements of the 3 greatest floods recorded at Beaucaire. 1840 and 1856 floods were recorded at Pont de Beaucaire gauge, where only daily measurements are available, but the peak stage is known. 2003 flood was recorded at Beaucaire Restitution gauge for which hourly measurements are available. The datum and the cross-section geometry of the two gauges are different, which explains the differences between the stages of the XIX\textsuperscript{th} and the XXI\textsuperscript{th} centuries}
	    \label{fig:FloodStages}
	\end{figure}

	\begin{figure}
		\centering
		\begin{subfigure}{.8\textwidth}
            \centering
            \includegraphics[width=\linewidth]{Supplementary/RC_Pt_zoom.pdf}
            \caption{}
            \label{subfig:RcPtZoom}
        \end{subfigure}
        \begin{subfigure}{0.8\textwidth}	
            \centering
            \includegraphics[width=\linewidth]{Supplementary/RC_Res_zoom.pdf}
            \caption{}
            \label{subfig:RcResZoom}
        \end{subfigure}
        \caption{Magnified rating curves of Pont de Beaucaire (a) and Beaucaire Restitution gauges (b) with 95\% uncertainty (in grey) with respect to maxpost values (solid lines). Dots with error bars represent the gaugings with 95\% uncertainty. Stable stage-discharge periods are numbered from the oldest to the latest (see table 5 in the article).}
	\end{figure}

	\begin{figure}[h]
		\centering
		\includegraphics[width=1\linewidth]{Supplementary/IC_AMAX_Both_Bands.pdf}
	    \caption{AMAX maxpost discharges (black solid line) time series with 95\% uncertainties from three different sources at Beaucaire (1816-2020). The three greatest floods are indicated. Vertical dotted lines represent rating shifts from table 5 and green bands represent the sub-samples used in section 4.5.3.}
	    \label{fig:AMAX_Bands}
	\end{figure}
